---
title: "National and Subnational estimates for Germany"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in Germany.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
---




```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 12, 
                      fig.width = 12,
                      dpi = 320)
```


```{r load-packages, include = FALSE}
library(EpiNow)
library(dplyr)
library(rmarkdown)
library(here)
library(stringr)
library(ggplot2)
library(sf)
```

## Expected daily cases by region

```{r}
summary_table <- readRDS("nowcast/regional-summary/summary_table.rds")
region_codes <- readRDS("nowcast/data/region_codes.rds")
latest_date <- readRDS("nowcast/regional-summary/latest_date.rds")

source("nowcast/utils/map_us.R")


summary_table <- summary_table %>% 
  dplyr::left_join(region_codes %>% 
                     dplyr::mutate(region = region %>% 
                                     stringr::str_replace("-", " ")), 
                   by = c("Region" = "region"))
```

```{r map, fig.width = 6, fig.height = 6, layout="l-body-outset", out.width = "80%"}
regions <- rnaturalearth::ne_states("Germany", returnclass = "sf")


regions_with_data <- regions %>%
   dplyr::left_join(summary_table,
                  by = c("iso_3166_2" = "region_code"))

map <- ggplot2::ggplot(regions_with_data) +
  ggplot2::geom_sf(ggplot2::aes(fill = `Expected change in daily cases`),
                   alpha = 0.8, size = 0.2, col = "white")


EpiNow::theme_map(
  map,
  continuous = FALSE,
  variable_label = "Expected change in daily cases",
  viridis_palette = "cividis",
  fill_labels = ggplot2::waiver()
  )
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated cases with a date of infection on the ", latest_date , ") in Germany, stratified by region, can be summarised by whether cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively.*")`

## National summary

```{r}
## Control parameters for region-report
summary_tables <- 0
summary_figures <- 1
title_depth <- 3
index <- 1
region <- "Germany"
region_path <- "_posts/global/nowcast/national"
show_title <- FALSE
```

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow")}
```



## Regional Breakdown

### Data availability

* Case counts by date, stratified by region were downloaded from a publically available source [@NCoVUtils; @gehrcke_covid-19_2020].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a publically available line-list of cases reported internationally [@kraemer2020epidemiological; @NCoVUtils].

### Limitations

* Line-list data to inform a region specific estimate of the reporting delay were not available. 

```{r} 
region_path <- NULL
summary_path <- "_posts/national/germany/nowcast/regional-summary"
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow")}
```
