---
title: "National and Subnational estimates for Canada"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in Canada. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Americas"
---


```{r load-custom-data}
library(magrittr)
region_codes <- covidregionaldata::get_regional_data(country = "Canada") %>% 
  dplyr::select(sovereignt = province, iso_3166_2) %>% 
  unique()
```

```{r load-geodata, include = FALSE}
geoData <- rnaturalearth::ne_states('Canada', returnclass = 'sf') 
geoData <- dplyr::left_join(geoData, region_codes, by = "iso_3166_2")
```


```{r load-defaults, include = FALSE}
region <- "Canada"
folder <- "canada"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
```


*Using data available up to the:* `r latest_date`

*Subnational estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/subnational/canada/cases/summary) and national estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/national/cases/summary).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*

## Interactive summary


```{r map, layout="l-body-outset", out.width = "200%", message = FALSE, warning = FALSE}
widget
```

<br>
`r widget_caption`

```{r,  child = here::here("utils", "national-template.Rmd")}
```

## Regional Breakdown

### Data availability

* Case counts by date, stratified by region, were downloaded from a public curated dataset [@covidregionaldata].
* Case onset dates were estimated using confirmed case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a Canadian specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.

```{r}
region_path <- NULL
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
