---
title: "Temporal variation in transmission during the COVID-19 outbreak"
site: distill::distill_website
bibliography: library.bib
---


```{r, include = FALSE}
title <- FALSE
```

```{r, child = "_paper/title-skeleton.Rmd"}
```



**More information** For more detail and the limitations of these estimates please see our [methods](https://epiforecasts.io/covid/methods.html)), our [preprint](https://wellcomeopenresearch.org/articles/5-112/v1)[@rtwebsite], the [documentation of our transmission tracking R package](https://epiforecasts.io/EpiNow2/)[@epinow2] and the documentation of our [R data package](https://epiforecasts.io/covidregionaldata/) [@covidregionaldata]. Stored versions of our estimates and the code used to generate them can be found [here](https://github.com/epiforecasts/covid-rt-estimates).**


**Note:** Our methodology remains an area of active research so please consider contacting us if interested in using, or evaluating, our approach. As an open-source project we also welcome code and methodology contributions, please see the linked code repositories for details on how to contribute.
