---
title: "National and Subnational estimates for Italy"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in Italy. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
  - "Italy"
---


```{r load-defaults, include = FALSE}
region <- "Italy"
folder <- "italy"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
source(here::here("utils/epi_leaflet.R"), local = TRUE)
```

```{r load-custom-data}
region_codes <- readRDS(here::here("_posts", "national", "italy", "region_codes.rds"))
region_codes <- dplyr::rename(region_codes, sovereignt = region, provnum_ne = region_code)
```

```{r load-geodata, include = FALSE}
geoData <- rnaturalearth::ne_states('Italy', returnclass = 'sf') %>%
  group_by(provnum_ne) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")
geoData <- dplyr::left_join(geoData, region_codes, by = "provnum_ne") %>%
  rename(Region = sovereignt)
```

*Using data available up to the:* `r latest_date`

*Subnational and national estimates are available to download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*

## National summary

```{r, include = FALSE}
summary_tables <- 0
summary_figures <- 0
title_depth <- 3
index <- 1
region_path <- "covid-rt-estimates/national/cases/national"
show_title <- FALSE
report_forecast <- TRUE
interactive <- TRUE
standalone <- FALSE
```

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow2")}
```

## Subnational breakdown

```{r,  child = here::here("utils", "national-map.Rmd")}
```

## Data availability

* Case counts by date, stratified by region, were constructed from daily data-sets made publicly available by the Dipartimento della Protezione Civile [@italydata; @covidregionaldata].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @covidregionaldata].

## Limitations

* Line-list data to inform a Italy specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.

