---
title: "National and Subnational estimates for Italy"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United States of America. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively (see Methods or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for further explanation).
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
---


```{r load-defaults, include = FALSE}
region <- "Italy"
folder <- "italy"
source(here::here("utils/subnational_defaults.R"))
```

```{r load-custom-data}
region_codes <- readRDS(here::here("_posts", "national", "italy", "region_codes.rds"))
```


*Using data available up to the:* `r latest_date`

*Subnational estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/subnational/italy/cases/summary) and national estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/national/cases/summary).*

## Expected daily confirmed cases by region

```{r map, fig.width = 6, fig.height = 6, layout="l-body-outset", out.width = "80%"}
summary_table <- summary_table %>%
  dplyr::rename(region = Region) %>%
  dplyr::left_join(region_codes, by = "region")

map <-  rnaturalearth::ne_states("Italy", returnclass = "sf") %>%
  dplyr::mutate(region_code = provnum_ne) %>% 
  dplyr::left_join(summary_table, by = "region_code") %>%
  ggplot2::ggplot(ggplot2::aes(fill = `Expected change in daily cases`)) +
  ggplot2::geom_sf(alpha = 0.8, size = 0.2)

map %>%
  EpiNow2::theme_map(
    continuous = FALSE,
    variable_label = "Expected change in daily cases",
    fill_labels = ggplot2::waiver(),
    breaks = levels(summary_table$`Expected change in daily cases`)
  )
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated cases with a date of infection on the ", latest_date , ") in Italy, stratified by region, can be summarised by whether cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively (see the [methods](https://epiforecasts.io/covid/methods.html) for details).*")`

## National summary

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow2")}
```


## Regional Breakdown

### Data availability

* Case counts by date, stratified by region, were constructed from daily data-sets made publicly available by the Dipartimento della Protezione Civile [@italydata; @covidregionaldata].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a Italy specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.

```{r}
region_path <- NULL
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
