---
title: "National and Subnational estimates for India"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in India. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Asia"
---

```{r load-geodata, include = FALSE}
geoData <- rnaturalearth::ne_states('India', returnclass = 'sf') 
geoData <-  dplyr::rename(geoData, Region = name_en)
```

```{r load-defaults, include = FALSE}
region <- "India"
folder <- "india"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
source(here::here("utils/epi_leaflet.R"), local = TRUE)
subnational_unit <- "State"
```

*Using data available up to the:* `r latest_date`

*Subnational and national estimates are available to download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*

```{r,  child = here::here("utils", "national-map.Rmd")}
```

### Data availability

* Case counts by date, stratified by region were downloaded from a publicly available source [@covidregionaldata; @noauthor_covid19-india_2020].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a publicly available line-list of cases reported internationally [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a region specific estimate of the reporting delay were not available.
