---
title: "National and Subnational estimates for India"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in India. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively (see Methods or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for further explanation).
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
---


```{r load-defaults, include = FALSE}
region <- "India"
folder <- "india"
source(here::here("utils/subnational_defaults.R"))
```

```{r load-custom-data}
region_codes <- readRDS(here::here("_posts", "national", "india", "region_codes.rds"))
```


*Using data available up to the:* `r latest_date`

*Subnational estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/subnational/india/cases/summary) and national estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/national/cases/summary).*

## Expected daily confirmed cases by region


```{r map, fig.height = 6, layout="l-body-outset", out.width = "70%"}
summary_table <- summary_table %>%
  dplyr::left_join(region_codes, by = c("Region" = "region"))

ladakh <- raster::getData('GADM', country='IND', level=2) %>%
  sf::st_as_sf() %>%
  dplyr::filter(NAME_2 == "Leh (Ladakh)" | NAME_2 == "Kargil") %>%
  dplyr::mutate(district = c("Leh", "Kargil"),
                region = "Ladakh") %>%
  dplyr::group_by(region) %>%
  dplyr::summarise()
regions <- raster::getData('GADM', country='IND', level=1) %>%
  sf::st_as_sf() %>%
  dplyr::select(region = NAME_1) %>%
  rbind(ladakh)
regions_with_data <- regions %>%
  dplyr::left_join(summary_table,
                   by = c("region" = "Region"))
map <- ggplot2::ggplot(regions_with_data) +
  ggplot2::geom_sf(ggplot2::aes(fill = `Expected change in daily cases`),
                   alpha = 0.8, size = 0.2, col = "white")
EpiNow2::theme_map(
  map,
  continuous = FALSE,
  variable_label = "Expected change in daily cases",
  fill_labels = ggplot2::waiver(),
  breaks = levels(summary_table$`Expected change in daily cases`)
  )
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated cases with a date of infection on the ", latest_date , ") in India, stratified by region, can be summarised by whether cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively (see the [methods](https://epiforecasts.io/covid/methods.html) for details). Regions with fewer than 40 cases reported on a single day are not included in the analysis (light grey).*")`

## National summary

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow2")}
```


## Regional Breakdown

### Data availability

* Case counts by date, stratified by region were downloaded from a publicly available source [@covidregionaldata; @noauthor_covid19-india_2020].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a publicly available line-list of cases reported internationally [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a region specific estimate of the reporting delay were not available.

```{r}
region_path <- NULL
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
