---
title: "National and Subnational estimates for the United States of America"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United States of America. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Americas"
---


```{r load-geodata, include = FALSE}
geoData <- rnaturalearth::ne_states('United States of America', returnclass = 'sf') 
geoData <- dplyr::mutate(geoData, sovereignt = name_en)
geoData <- dplyr::filter(geoData, !sovereignt %in% c("Alaska", "Hawaii"))
```

```{r load-defaults, include = FALSE}
region <- "United States"
folder <- "united-states"
region_var <- "state"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
```

*Using data available up to the:* `r latest_date`

*Subnational estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/subnational/united-states/cases/summary) and national estimates are available to download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/national/cases/summary).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*

## Interactive summary

```{r map, layout="l-body-outset", out.width = "200%", message = FALSE, warning = FALSE}
widget
```

<br>
`r widget_caption`

```{r,  child = here::here("utils", "national-template.Rmd")}
```

## Regional Breakdown

* Case counts by date, stratified by region, were downloaded from a publically available source  curated by Johns Hopkins University [@jhu_2019_2020; @covidregionaldata].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a publically available line-list of international cases [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a State specific estimates of the reporting delay were not available.

```{r}
region_path <- NULL
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
