---
title: "National and Subnational estimates for the United States of America"
description: |
  Estimates presented here are no longer updated as of 31 March 2022. For more information on the rationale behind this move and some reflections on 2 years of global nowcasting and forecasting, please read our related [blog post](https://epiforecasts.io/posts/2022-03-25-rt-reflections/).
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Americas"
  - "United States"
---


```{r load-geodata, include = FALSE}
library(magrittr)
geoData <- rnaturalearth::ne_states('United States of America', returnclass = 'sf') 
geoData <- dplyr::mutate(geoData, sovereignt = name_en)
geoData <- dplyr::filter(geoData, !sovereignt %in% c("Alaska", "Hawaii")) %>%
  dplyr::rename(Region = sovereignt)
```

```{r load-defaults, include = FALSE}
region <- "United States"
folder <- "united-states"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
source(here::here("utils/epi_leaflet.R"), local = TRUE)
subnational_unit <- "State"
summary_table <- summary_table %>%
  rename(Region = State)
```

*Using data available up to the:* `r latest_date`

*Subnational and national estimates are available to download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*

## National summary

```{r, include = FALSE}
summary_tables <- 0
summary_figures <- 0
title_depth <- 3
index <- 1
region_path <- "covid-rt-estimates/national/cases/national"
show_title <- FALSE
report_forecast <- TRUE
interactive <- TRUE
standalone <- FALSE
```

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow2")}
```

## Subnational breakdown

```{r}
fig_tab_no <- 2
```

```{r,  child = here::here("utils", "national-map.Rmd")}
```

### Data availability

* Case counts by date, stratified by region, were downloaded from a publically available source  curated by Johns Hopkins University [@jhu_2019_2020; @covidregionaldata].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a publically available line-list of international cases [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a State specific estimates of the reporting delay were not available.
