---
title: "National and Subnational estimates for the United States of America"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United States of America.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
---




```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 12, 
                      fig.width = 12,
                      dpi = 320)
```


```{r load-packages, include = FALSE}
library(EpiNow)
library(dplyr)
library(rmarkdown)
library(here)
library(stringr)
library(ggplot2)
library(sf)
```

## Expected daily cases by region

```{r}
summary_table <- readRDS("nowcast/regional-summary/summary_table.rds")
region_codes <- readRDS("nowcast/data/region_codes.rds")
latest_date <- readRDS("nowcast/regional-summary/latest_date.rds")
source("nowcast/utils/map_us.R")


summary_table <- summary_table %>% 
  dplyr::left_join(region_codes, by = c("State" = "region"))
```

```{r map, fig.width = 12, fig.height = 6, layout="l-body-outset"}
map_us(summary_table,
              custom_theme = function(.) {
    EpiNow::theme_map(
    .,
    continuous = FALSE,
    variable_label = "Expected change in daily cases",
    viridis_palette = "cividis",
    fill_labels = ggplot2::waiver()
  )})
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated cases with a date of infection on the ", latest_date , ") in the United States of America, stratified by state, can be summarised by whether cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively.*")`

## National summary

```{r}
## Control parameters for region-report
summary_tables <- 0
summary_figures <- 1
title_depth <- 3
index <- 1
region <- "United States of America"
region_path <- "_posts/global/nowcast/national"
show_title <- FALSE
```

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow")}
```


## Regional Breakdown

### Data availability

* Case counts by date, stratified by region, were downloaded from a publically available source  curated by Johns Hopkins University [@jhu_2019_2020; @NCoVUtils].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a publically available line-list of international cases [@kraemer2020epidemiological; @NCoVUtils].

### Limitations

* Line-list data to inform a State specific estimates of the reporting delay were not available. 


```{r} 
region_path <- NULL
summary_path <- "_posts/national/united-states/nowcast/regional-summary"
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow")}
```

