---
title: "National and Subnational estimates for Italy"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in Italy.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 3
---




```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      dpi = 320)
```


```{r load-packages, include = FALSE}
library(EpiNow)
library(dplyr)
library(rmarkdown)
library(here)
```

## Expected daily cases by region

```{r}
summary_table <- readRDS ("nowcast/summary/summary_table.rds")
region_codes <- readRDS("nowcast/data/region_codes.rds")
```

```{r map, fig.width = 6, fig.height = 6, layout="l-body-outset"}
summary_table %>% 
  dplyr::rename(region  = Region) %>% 
  dplyr::left_join(region_codes, by = "region") %>% 
  EpiNow::country_map(country = "Italy", variable = "Expected change in daily cases") +
  ggplot2::guides(fill = ggplot2::guide_legend(title = "Expected change in daily cases",
                                               ncol = 2))
```

<br>
*Figure 1: Map of the expected change in daily cases*

## National summary

```{r}
## Control parameters for region-report
summary_tables <- 0
summary_figures <- 1
title_depth <- 3
index <- 1
region <- "italy"
path <- "_posts/global"
show_title <- FALSE
```

```{r,  child = "../../templates/_region-report.Rmd"}
```


## Regional Breakdown

### Data availability

* Case counts by date, stratified by region, were constructed from daily datasets made publically available by the Dipartimento della Protezione Civile [@italydata; @NCoVUtils].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to a European line-list [@kraemer2020epidemiological; @NCoVUtils].

### Limitations

* Line-list data to inform a Italy specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.


```{r} 
region_reports <- TRUE
path <- "_posts/italy"
```

```{r,  child = "../../templates/_regional-summary.Rmd"}
```

