---
title: "National and Subnational estimates for the United Kingdom"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United Kingdom. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively (see Methods for further explanation).
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
---




```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 12, 
                      fig.width = 12,
                      dpi = 320)
```


```{r load-packages, include = FALSE}
library(EpiNow)
library(dplyr)
library(rmarkdown)
library(here)
```

## Expected daily cases by region

```{r}
summary_table <- readRDS ("nowcast/regional-summary/summary_table.rds")
uk_shp <- readRDS("nowcast/data/uk_shp.rds")
latest_date <- readRDS("nowcast/regional-summary/latest_date.rds")
```

```{r map, fig.width = 6, fig.height = 6, layout="l-body-outset"}
map <- uk_shp %>% 
  dplyr::full_join(summary_table %>% 
  dplyr::rename(region  = Region), by = "region") %>% 
  ggplot2::ggplot(ggplot2::aes(fill = `Expected change in daily cases`)) +
  ggplot2::geom_sf(alpha = 0.8, size = 0.2)

map %>% 
  EpiNow::theme_map(
    continuous = FALSE,
    variable_label = "Expected change in daily cases",
    viridis_palette = "cividis",
    fill_labels = ggplot2::waiver()
  )
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated cases with a date of infection on the ", latest_date , ") in the United Kingdom, stratified by region, can be summarised by whether cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively.*")`


## National summary

```{r}
## Control parameters for region-report
summary_tables <- 0
summary_figures <- 1
title_depth <- 3
index <- 1
region <- "United Kingdom"
region_path <- "_posts/global/nowcast/national"
show_title <- FALSE
```

```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow")}
```


## Regional Breakdown

### Data availability

* Case counts by date, stratified by region, were downloaded from a public curated dataset [@doughty_visualisation_2020; @NCoVUtils].
* Case onset dates were estimated using case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @NCoVUtils].

### Limitations

* Line-list data to inform a United Kingdom specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.
* Regions are only included if at least 40 cases have been reported on a single day.


```{r} 
standalone <- FALSE
region_path <- NULL
summary_path <- "_posts/national/united-kingdom/nowcast/regional-summary"
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow")}
```

