---
title: "National and Subnational estimates for the United Kingdom"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United Kingdom. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively (see Methods or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for further explanation).
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
---


```{r load-defaults, include = FALSE}
region <- "United Kingdom"
folder <- "united-kingdom"
source(here::here("utils/subnational_defaults.R"))
```

```{r load-custom-data}
uk_shp <- readRDS(here::here("_posts", "national", "united-kingdom", "uk_shp.rds"))
```

*Using data available up to the:* `r latest_date`

*These estimates are available for download [here](https://github.com/epiforecasts/covid-rt-estimates/tree/master/subnational/united-kingdom/cases/summary).*


## Expected daily confirmed cases by region

```{r map,fig.height = 8, layout="l-body-outset", out.width = "70%", warning = FALSE, message = FALSE}
map <- uk_shp %>%
  dplyr::rename(region = nuts118nm) %>% 
  dplyr::left_join(summary_table %>%
  dplyr::rename(region  = Region), by = "region") %>%
  ggplot2::ggplot(ggplot2::aes(fill = `Expected change in daily cases`)) +
  ggplot2::geom_sf(alpha = 0.8, size = 0.2)

map %>%
  EpiNow2::theme_map(
    continuous = FALSE,
    variable_label = "Expected change in daily cases",
    fill_labels = ggplot2::waiver(),
    breaks = levels(summary_table$`Expected change in daily cases`)
  )
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated confirmed cases with a date of infection on the ", latest_date , ") in the United Kingdom, stratified by region, can be summarised by whether confirmed cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively (see the [methods](https://epiforecasts.io/covid/methods.html) for details).*")`


## National summary


```{r,  child = system.file("templates/_region-report.Rmd", package = "EpiNow2")}
```

## Regional Breakdown

### Data availability

* Case counts by date, stratified by region, were downloaded from a public curated dataset [@noauthor_coronavirus_2020; @white_coronavirus_2020; @covidregionaldata].
* Case onset dates were estimated using confirmed case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

* Line-list data to inform a United Kingdom specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.

```{r}
region_path <- NULL
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
