---
title: "National and Subnational estimates for the United Kingdom"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United Kingdom. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
author:
  - name: Katharine Sherratt*, Sam Abbott*, Sophie R Meakin, Joel Hellewell, James D Munday, Nikos Bosse, CMMID Covid-19 working group, Mark Jit, Sebastian Funk
    affiliation: "Centre for Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine"
date: "2020-10-20"
journal: "medRxiv preprint"
doi: "10.1101/2020.10.18.20214585"
citation_url: https://www.medrxiv.org/content/10.1101/2020.10.18.20214585v1
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
---

```{r load-geodata, include = FALSE}
geoData <- rnaturalearth::ne_states('United Kingdom', returnclass = 'sf')
geoData <- dplyr::mutate(geoData, sovereignt = geonunit)
```

```{r load-defaults, include = FALSE}
library(patchwork)
library(RColorBrewer)
library(here)
library(ggplot2)
library(dplyr)
library(readr)
library(cowplot)
library(data.table)
library(forecast)
library(knitr)
library(kableExtra)
library(DT)
library(tidyr)
library(stringr)
region <- "United Kingdom"
folder <- "united-kingdom"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
```

*Using data available up to the:* `r latest_date`

*Subnational and national estimates are available to download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived, and a [pre-print](https://www.medrxiv.org/content/10.1101/2020.10.18.20214585v1) for methods comparing Rt estimates by data source*

<!-- ## Interactive summary -->

```{r map, layout="l-body-outset", out.width = "200%", message = FALSE, warning = FALSE, eval = FALSE}
widget
```

<!-- <br> `r widget_caption` -->

```{r,  child = here::here("utils", "national-template.Rmd")}
```

## Comparing estimates from cases, admissions, and deaths data

We calculate Rt as the average of how many new infections arise from one infected person. However, each data source for SARS-CoV-2 infections (test-positive cases, hospital admissions, or deaths) represents a slightly different type of "average" person who has been infected. For example, when most new infections are spread between young people who may be less vulnerable to severe disease, Rt calculated from all test-positive cases might be expected to be greater than an Rt calculated from patients in hospital who would, generally, represent older individuals more at risk of having severe outcomes from SARS-CoV-2 infection. 

All the Rt estimates are referenced to the same date, the estimated date at which people who end up being reported as cases, admissions or deaths became infected. Therefore, if all the data streams represented the same patterns of spread the lines from the three colours would be expected to completely overlap. Where they do not, this tells us something about how Covid-19 is spreading in populations of different levels of vulnerability populations. To explore in more depth how Rt from different data sources can be used to understand transmission dynamics across the population, and a more detailed methodology, see our pre-print [@rt-comparison].

The latest date shown for estimates is based on the most recent date that we believe to have a reliable count. In the UK, this lags behind the current date by several days. This is because of the delay in reporting a new infection. For example, a positive test can take up to five days to be reported in public data from the date on which it was taken, meaning that data less than five days old will not include all positive tests for that day. Moreover, a test would usually be done following onset of symptoms, and this usually happens several days after initial infection. The reporting also varies by data source, where we account for a longer delay in reporting deaths compared to hospital admissions or positive tests.

```{r rt-comparison-nations, fig.height = 12, fig.width = 14, out.width = "95%", echo=FALSE, message = FALSE, warning = FALSE}
data_path <- here("covid-rt-estimates/subnational/united-kingdom")
# Rt
summary <- fread(file.path(data_path, "collated/rt/summary_latest.csv"))
summary <- mutate(summary, 
                  'Data source' = dplyr::recode_factor(source,
                                                       "cases" = "Test-positives",
                                                       "admissions" = "Hospital admissions",
                                                       "deaths" = "Deaths"))
# Data
admissions <- fread(file.path(data_path, "admissions/summary/reported_cases.csv"))
deaths <- fread(file.path(data_path, "deaths/summary/reported_cases.csv"))
test_positive <- fread(file.path(data_path, "cases/summary/reported_cases.csv"))

data <- rbindlist(list("admissions" = admissions, 
                       "deaths" = deaths, 
                       "cases" = test_positive), idcol = "source")
setnames(data, "confirm", "count")
data <- data[, count := as.numeric(count)]
data <- data[, count_ma := ma(as.numeric(count), 7), by = .(source, region)]
# Join
summary <- merge(summary, data, by = c("region", "date", "source"), all.x = TRUE)

# Plotting features
colours <- brewer.pal(name = "Dark2", n = 6)[c(1, 4, 6)]
names(colours) <- c("Test-positives", "Hospital admissions", "Deaths")

region_names <- factor(c("England", "Scotland", "Wales", "Northern Ireland",
                  "North East and Yorkshire", "North West", "Midlands", "East of England",
                  "London", "South East", "South West"), ordered = TRUE)

plot_rt <- function(region_name) {
  ggplot(summary[region %in% region_name], aes(x = date, col = `Data source`, fill = `Data source`)) +
      geom_ribbon(aes(ymin = lower_20, ymax = upper_20), alpha = 0.6, size = 0, colour = NA) +
      geom_ribbon(ggplot2::aes(ymin = lower_50, ymax = upper_50), alpha = 0.4, size = 0, colour = NA) +
      geom_ribbon(ggplot2::aes(ymin = lower_90, ymax = upper_90), alpha = 0.2, size = 0.05) +
      geom_hline(yintercept = 1, linetype = 2) +
      labs(y = "R(t)", x = NULL) +
      scale_color_manual(values = colours) +
      scale_fill_manual(values = colours) +
      scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
      theme_cowplot() +
      theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1)) +
      guides(colour = ggplot2::guide_legend(override.aes = list(alpha = 1)))
}
region_plot_data <- purrr::map(region_names, ~ plot_rt(region_name = .x))

plot_cases <- function(region_name) {
    ggplot(summary[region %in% region_name], aes(x = date, col = `Data source`, fill = `Data source`)) +
    geom_point(ggplot2::aes(y = count)) +
    geom_line(ggplot2::aes(y = count_ma)) +
    labs(y = "Log count", x = NULL, title = region_name) +
    scale_color_manual(values = colours) +
    scale_fill_manual(values = colours) +
    scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
    theme_cowplot() +
    theme(axis.text.x = ggplot2::element_blank()) +
    guides(colour = "none", fill = "none") +
    scale_y_log10(labels = scales::label_comma(accuracy = 1))
}
region_plot_rt <- purrr::map(region_names,  ~ plot_cases(region_name = .x))

# # Plot selected regions:
((region_plot_data[[1]] / region_plot_rt[[1]]) |
  (region_plot_data[[2]] / region_plot_rt[[2]])) /
  ((region_plot_data[[3]] / region_plot_rt[[3]]) |
  (region_plot_data[[4]] / region_plot_rt[[4]])) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```
*Figure 2: Raw data counts with 7-day moving average; and estimates of Rt (lightest ribbon = 90% credible interval; darker ribbon = the 50% credible interval, darkest ribbon = 20% credible interval). Rt derived from data sources including all test-positive cases, hospital admissions, and deaths with a positive test in the previous 28 days for the devolved authorities of the United Kingdom.*

```{r rt-comparison-regions, fig.height = 18, fig.width = 14, out.width = "95%", echo = FALSE, message = FALSE, warning = FALSE}
((region_plot_data[[5]] / region_plot_rt[[5]]) |
  (region_plot_data[[6]] / region_plot_rt[[6]])) /

  ((region_plot_data[[7]] / region_plot_rt[[7]]) |
  (region_plot_data[[8]] / region_plot_rt[[8]])) /

  ((region_plot_data[[9]] / region_plot_rt[[9]]) |
  (region_plot_data[[10]] / region_plot_rt[[10]])) /

  ((region_plot_data[[11]] / region_plot_rt[[11]])  |
  (plot_spacer() / plot_spacer())) +

plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

*Figure 3: Raw data counts with 7-day moving average; and estimates of Rt (lightest ribbon = 90% credible interval; darker ribbon = the 50% credible interval, darkest ribbon = 20% credible interval). Rt derived from data sources including all test-positive cases, hospital admissions, and deaths with a positive test in the previous 28 days in the NHS regions of England.*

```{r fig-caption-dates, echo = FALSE}
collated_table <- summary %>%
  dplyr::group_by(region, source) %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::ungroup()  
latest_date_cases <- collated_table %>%
  dplyr::filter(source == "cases") %>%
  dplyr::pull(date) %>%
  max()
latest_date_admissions <- collated_table %>%
  dplyr::filter(source == "admissions") %>%
  dplyr::pull(date) %>%
  max()
latest_date_deaths <- collated_table %>%
  dplyr::filter(source == "deaths") %>%
  dplyr::pull(date) %>%
  max()
```

*Table 2: Latest estimates of Rt, derived from data sources including all test-positive cases, hospital admissions, and deaths with a positive test in the previous 28 days. The median and 90% credible interval is shown. Latest dates vary for cases (`r latest_date_cases`), admissions (`r latest_date_admissions`), and deaths (`r latest_date_deaths`).*

```{r, echo = FALSE}
collated_table <- collated_table %>%
  mutate(combine = str_c(median, " (", lower_90, " -- ", upper_90, ")")) %>%
  select("Data source", region, combine) %>%
  pivot_wider(names_from = "Data source", values_from = "combine") %>%
  rename(Region = region)
if (!interactive) {
  collated_table %>%
    kable(booktabs = TRUE, longtable = TRUE) %>%
    landscape()
} else {
  collated_table %>%
    datatable(rownames = FALSE,
              extensions = c("Buttons"),
              options = list(dom = "Bfrtip", buttons = c("csv")))
}
```

## Regional Breakdown

Estimates are based on test positive cases by specimen date for NHS regions in England and by date of report in Wales, Scotland and Northern Ireland.

### Data availability

-   Case counts by date, stratified by region, were downloaded from a public curated dataset [@noauthor_coronavirus_2020; @white_coronavirus_2020; @covidregionaldata].
-   Case onset dates were estimated using confirmed case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

- Line-list data to inform a United Kingdom specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.

```{r}
region_path <- NULL
case_def <- "case"
folder <- "united-kingdom"
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
fig_start <- 4
tab_start <- 3
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
