---
title: "National and Subnational estimates for the United Kingdom"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally in the United Kingdom. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
author:
  - name: Katharine Sherratt*, Sam Abbott*, Sophie R Meakin, Joel Hellewell, James D Munday, Nikos Bosse, CMMID Covid-19 working group, Mark Jit, Sebastian Funk
    affiliation: "Centre for Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine"
date: "2020-10-20"
journal: "medRxiv preprint"
doi: "10.1101/2020.10.18.20214585"
citation_url: https://www.medrxiv.org/content/10.1101/2020.10.18.20214585v1
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Subnational estimates"
  - "Europe"
---

```{r load-geodata, include = FALSE}
geoData <- rnaturalearth::ne_states('United Kingdom', returnclass = 'sf') 
geoData <- dplyr::mutate(geoData, sovereignt = geonunit)
```

```{r load-defaults, include = FALSE}
library(patchwork)
region <- "United Kingdom"
folder <- "united-kingdom"
source(here::here("utils/subnational_defaults.R"), local = TRUE)
```

*Using data available up to the:* `r latest_date`

*Subnational and national estimates are available to download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived, and a [pre-print](https://www.medrxiv.org/content/10.1101/2020.10.18.20214585v1) for methods comparing Rt estimates by data source*

<!-- ## Interactive summary -->

```{r map, layout="l-body-outset", out.width = "200%", message = FALSE, warning = FALSE, eval = FALSE}
widget
```

<!-- <br> `r widget_caption` -->

```{r,  child = here::here("utils", "national-template.Rmd")}
```

## Comparing estimates from cases, admissions, and deaths data

We calculate Rt as the average of how many new infections arise from one infected person. However, each data source for Covid-19 infections (test-positive cases, hospital admissions, or deaths) represents a slightly different type of "average" person who has been infected. For example, when most new infections are spread between young people who may be less vulnerable to severe disease, the Rt from all test-positive cases rises quicker than the Rt calculated from patients in hospital. This happened in the UK in August. This comparison can help us track in real-time how Covid-19 is spreading in the most vulnerable populations. 

To explore in more depth how Rt from different data sources can be used to understand transmission dynamics across the population, and a more detailed methodology, see our pre-print [@rt-comparison].

```{r rt-comparison-nations, fig.height = 12, fig.width = 14, out.width = "95%", echo=FALSE}
# Rt
summary <- suppressMessages(readr::read_csv(here::here("covid-rt-estimates/subnational/united-kingdom/collated/rt/summary_latest.csv"))) %>%
  dplyr::mutate('Data source' = dplyr::recode_factor(source, 
                                                     "cases" = "Test-positives",
                                                     "admissions" = "Hospital admissions",
                                                     "deaths" = "Deaths"))
# Data
data <- suppressMessages(covidregionaldata::get_regional_data("uk", nhsregions = TRUE)) %>%
  dplyr::filter(date %in% summary$date) %>%
  dplyr::select(date, region, 
                admissions = hosp_new_blend, 
                cases = cases_new, deaths = deaths_new) %>%
  tidyr::pivot_longer(cols = c(admissions, cases, deaths), 
                      names_to = "source", values_to = "count") %>%
  dplyr::group_by(source, region) %>%
  dplyr::mutate(count_ma = forecast::ma(count, 7))

# Join
summary <- dplyr::left_join(summary, data, by = c("region", "date", "source"))

# Plotting features
colours <- c("Test-positives" = "#1b9e77","Hospital admissions" =  "#7570b3","Deaths" = "#d95f02")
region_names <- factor(c("England", "Scotland", "Wales", "Northern Ireland",
                  "North East and Yorkshire", "North West", "Midlands", "East of England", 
                  "London", "South East", "South West"), ordered = TRUE)
# Rt plot
plot_fn <- function(region_name, rt = TRUE){
  
  region_summary <- summary %>%
    dplyr::filter(region %in% region_name) 
  
  p <- region_summary %>%
    ggplot2::ggplot(ggplot2::aes(x = date, col = `Data source`, fill = `Data source`)) 
  
  if (rt) {
  p + 
    ggplot2::geom_ribbon(ggplot2::aes(ymin = lower_20, ymax = upper_20),
                alpha = 0.2, size = 0, colour = NA) +  
    ggplot2::geom_ribbon(ggplot2::aes(ymin = lower_50, ymax = upper_50),
                alpha = 0.1, size = 0, colour = NA) +
    ggplot2::geom_hline(yintercept = 1, linetype = 2) +
    ggplot2::labs(y = "R(t)", x = NULL) +
    ggplot2::scale_color_manual(values = colours) +
    ggplot2::scale_fill_manual(values = colours) +
    ggplot2::scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
    cowplot::theme_cowplot() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1)) +
    ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(alpha = 1)))
  }
  
    else {
  p +
    ggplot2::geom_point(ggplot2::aes(y = count)) +
    ggplot2::geom_line(ggplot2::aes(y = count_ma)) +
    ggplot2::labs(y = "Log count", x = NULL, title = region_name) +
    ggplot2::scale_color_manual(values = colours) +
    ggplot2::scale_fill_manual(values = colours) +
    ggplot2::scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
    cowplot::theme_cowplot() +
    ggplot2::theme(axis.text.x = ggplot2::element_blank()) +
    ggplot2::guides(colour = "none", fill = "none") +
    ggplot2::scale_y_log10()
  }
}

region_plot_data <- purrr::map(region_names,
                             ~ plot_fn(region_name = .x, rt = FALSE))
region_plot_rt <- purrr::map(region_names,
                             ~ plot_fn(region_name = .x, rt = TRUE))

# # Plot selected regions:
((region_plot_data[[1]] / region_plot_rt[[1]]) | 
  (region_plot_data[[2]] / region_plot_rt[[2]])) /
  ((region_plot_data[[3]] / region_plot_rt[[3]]) | 
  (region_plot_data[[4]] / region_plot_rt[[4]])) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "bottom")

```
*Figure 2: Raw data counts with 7-day moving average; and estimates of Rt (median, with 50% (darker shade) and 90% (lightest shade) credible interval), derived from data sources including all test-positive cases, hospital admissions, and deaths with a positive test in the previous 28 days for the devolved authorities of the United Kingdom.*

```{r rt-comparison-regions, fig.height = 18, fig.width = 14, out.width = "95%", echo = FALSE}
((region_plot_data[[5]] / region_plot_rt[[5]]) | 
  (region_plot_data[[6]] / region_plot_rt[[6]])) /
  
  ((region_plot_data[[7]] / region_plot_rt[[7]]) | 
  (region_plot_data[[8]] / region_plot_rt[[8]])) /
  
  ((region_plot_data[[9]] / region_plot_rt[[9]]) | 
  (region_plot_data[[10]] / region_plot_rt[[10]])) /
  
  ((region_plot_data[[11]] / region_plot_rt[[11]])  |
  (patchwork::plot_spacer() / patchwork::plot_spacer())) +
  
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "bottom")

```

*Figure 3: Raw data counts with 7-day moving average; and estimates of Rt (median, with 50% (darker shade) and 90% (lightest shade) credible interval), derived from data sources including all test-positive cases, hospital admissions, and deaths with a positive test in the previous 28 days in the NHS regions of England.*

```{r fig-caption-dates, echo = FALSE}
collated_table <- summary %>%
  dplyr::group_by(region, source) %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::ungroup()  
latest_date_cases <- collated_table %>%
  dplyr::filter(source == "cases") %>%
  dplyr::pull(date) %>%
  max()
latest_date_admissions <- collated_table %>%
  dplyr::filter(source == "admissions") %>%
  dplyr::pull(date) %>%
  max()
latest_date_deaths <- collated_table %>%
  dplyr::filter(source == "deaths") %>%
  dplyr::pull(date) %>%
  max()
```

*Table 2: Latest estimates of Rt, derived from data sources including all test-positive cases, hospital admissions, and deaths with a positive test in the previous 28 days. The median and 90% credible interval is shown. Latest dates vary for cases (`r latest_date_cases`), admissions (`r latest_date_admissions`), and deaths (`r latest_date_deaths`).*

```{r, echo = FALSE}
collated_table <- collated_table %>%
  dplyr::mutate(combine = stringr::str_c(median, " (", lower_90, " -- ", upper_90, ")")) %>%
  dplyr::select("Data source", region, combine) %>%
  tidyr::pivot_wider(names_from = "Data source", values_from = "combine") %>%
  dplyr::rename(Region = region)
if (!interactive) {
   collated_table %>% 
   knitr::kable(booktabs = TRUE, longtable = TRUE) %>% 
   kableExtra::landscape()
  
} else {
  collated_table %>% 
    DT::datatable(rownames = FALSE,
                  extensions = c("Buttons"), 
                  options = list(dom = "Bfrtip", buttons = c("csv")))
} 
```

## Regional Breakdown

### Data availability

-   Case counts by date, stratified by region, were downloaded from a public curated dataset [@noauthor_coronavirus_2020; @white_coronavirus_2020; @covidregionaldata].
-   Case onset dates were estimated using confirmed case counts by date of report and a distribution of reporting delays fitted to an international line-list [@kraemer2020epidemiological; @covidregionaldata].

### Limitations

-   Line-list data to inform a United Kingdom specific estimate of the reporting delay was not available. This means that we could also not account for any regional differences.

```{r}
region_path <- NULL
case_def <- "case"
folder <- "united-kingdom"
summary_path <- file.path("covid-rt-estimates/subnational", folder, "cases/summary")
fig_start <- 4
tab_start <- 3
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```