---
title: "Global summary"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      fig.width = 12, fig.height = 12,
                      message = FALSE,
                      warning = FALSE,
                      dpi = 320)
```


```{r load-packages, include = FALSE}
library(EpiNow)
library(dplyr)
library(rmarkdown)
library(here)
```

### Expected daily cases by country

```{r}
summary_table <- readRDS("nowcast/national-summary/summary_table.rds")
latest_date <- readRDS("nowcast/national-summary/latest_date.rds")
```

```{r map, fig.width = 12, fig.height = 6, layout="l-body-outset"}
summary_table %>% 
  dplyr::rename(country = `Country/Region`) %>% 
  dplyr::filter(!country %in% c("Hubei", "Hong Kong", "China excluding Hubei")) %>% 
  EpiNow::global_map(variable = "Expected change in daily cases")
```

<br>
`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated cases with a date of infection on the ", latest_date , ") can be summarised by whether cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively. Countries are only included if at least 100 cases have been reported on a single day.*")`


```{r} 
standalone <- TRUE
region_path <- NULL
summary_path <- "_posts/global/nowcast/national-summary"
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow")}
```
