---
title: "Global summary"
description: |
  Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting both nationally and subnationally. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
bibliography: library.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_depth: 2
categories:
  - "Global summary"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      fig.width = 12, fig.height = 12,
                      message = FALSE,
                      warning = FALSE,
                      dpi = 400)
```

```{r load-packages, include = FALSE}
library(EpiNow2)
library(RtD3)
library(dplyr)
library(magrittr)
library(rmarkdown)
library(here)
source(here::here("utils/epi_leaflet.R"), local = TRUE)
```

```{r}
summary_table <- data.table::fread(here::here("covid-rt-estimates", "national", "cases",
                                    "summary", "summary_table.csv"))
summary_table[, data := "Test positive case notifications"]

deaths_table <- data.table::fread(here::here("covid-rt-estimates", "national", "deaths",
                                    "summary", "summary_table.csv"))
deaths_table[, data := "Deaths attributed to Covid-19"]
overall_table <- rbind(summary_table, deaths_table)
latest_date <- readRDS(here::here("covid-rt-estimates", "national", "cases",
                                    "summary", "latest_date.rds"))
```

*Using data available up to the:* `r latest_date`

*Estimates (based on reported cases and deaths) are available for download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*

```{r map, fig.height=0.9,layout="l-body-outset", out.width = "100%", warning = FALSE, message = FALSE}
geoData <- rnaturalearth::ne_countries(scale = "medium", returnclass = 'sf')

geoData <- geoData %>%
  dplyr::mutate(Country = ifelse(sovereignt == "United States of America",
                                 "United States", sovereignt),
                Country = ifelse(Country == "Ivory Coast",
                                 "Côte d’Ivoire", Country),
                Country = ifelse(Country == "Democratic Republic of the Congo",
                                 "Congo - Kinshasa", Country),
                Country = ifelse(Country == "Republic of the Congo",
                                 "Congo - Brazzaville", Country),
                Country = ifelse(Country == "Bosnia and Herzegovina",
                                 "Bosnia & Herzegovina", Country),
                Country = ifelse(Country == "Macedonia",
                                 "North Macedonia", Country),
                Country = ifelse(Country == "Myanmar",
                                 "Myanmar (Burma)", Country)
                )
geoData %>%
  mutate(link = paste0("https://epiforecasts.io/covid/posts/national/",
                       sub(" ", "-", tolower(Country)))) %>% 
  epi_leaflet(overall_table, label_style = "popup", link_text = "Details")
```

`r paste0("*Figure 1: The results of the latest reproduction number estimates (based on estimated confirmed cases with a date of infection on the ", latest_date , ") can be summarised by whether confirmed cases are likely increasing or decreasing. This represents the strength of the evidence that the reproduction number in each region is greater than or less than 1, respectively (see the [methods](https://epiforecasts.io/covid/methods.html) for details). Click on a country (or search) to see national level estimates (and a link to more detailed estimates.*")`

```{r}
tab_start <- 1
fig_start <- 1
case_def <- "case"
summary_path <- "covid-rt-estimates/national/cases/summary"
interactive <- TRUE
standalone <- FALSE
report_forecast <- TRUE
region_path <- NULL
all_regions <- FALSE
```

```{r,  child = system.file("templates/_regional-summary.Rmd", package = "EpiNow2")}
```
