---
title: "Estimates for Isle of Man"
description: |
 Identifying changes in the reproduction number, rate of spread, and doubling time during the course of the COVID-19 outbreak whilst accounting for potential biases due to delays in case reporting in Isle of Man. These results are impacted by changes in testing effort, increases and decreases in testing effort will increase and decrease reproduction number estimates respectively.
bibliography: library.bib
output:
 distill::distill_article:
   self_contained: true
   toc: true
   toc_depth: 2
categories:
 - 'Europe'
---


*Using data available up to the:* `r Sys.Date()`

*Note that it takes time for infection to cause symptoms, to get tested for SARS-CoV-2 infection, for a positive test to return and ultimately to enter the case data presented here. In other words, today’s case data are only informative of new infections about two weeks ago. This is reflected in the plots below, which are by date of infection.*

*Estimates are available to download [here](https://dataverse.harvard.edu/dataverse/covid-rt).*

*See our see [Methods](https://epiforecasts.io/covid/methods) or our [paper](https://wellcomeopenresearch.org/articles/5-112/) for an explanation of how these estimates are derived.*


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 12,
                      fig.width = 12,
                      dpi = 300)
```



```{r load-packages, include = FALSE}
library(EpiNow2)
library(data.table)
library(dplyr)
library(magrittr)
library(rmarkdown)
library(here)

latest_date <- readRDS(here::here('covid-rt-estimates/national/cases/summary/latest_date.rds'))
```




```{r }
## Control parameters for region-report
summary_tables <- 0
summary_figures <- 0
title_depth <- 3
index <- 1
region <- "Isle of Man"
region_path <- "covid-rt-estimates/national/cases/national"
show_title <- FALSE
report_forecast <- TRUE
```




```{r child = system.file('templates/_region-report.Rmd', package = 'EpiNow2')}
```

